---
output:
  rmdformats::html_clean:
    highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Data visualization - independent exercises

#### **1. Getting the data**

**1.1** Just like we did yesterday, install the package `rjstat` (`install.packages("rjstat")`) if you do not have it already installed. Then load the necessary libraries by copy-pasting the following lines into an R script and running them in RStudio.

```{r}
library(rjstat)
library(tidyverse)
```

**1.2** Retrieve the coronavirus observations data from the THL website. We will use the same data that we worked with yesterday.

Getting the data was covered in the Day 1 extra exercises, so you can run through the following code chunk in your Rstudio session. You can highlight the code below and copy-paste it into a new R script.

```{r}

# Getting the data
url <- "https://sampo.thl.fi/pivot/prod/en/epirapo/covid19case/fact_epirapo_covid19case.json?"

measure.infections <- "measure-444833"
time.day <- "dateweek2020010120201231-443702L"
region.shp <- "hcdmunicipality2020-445222"

# Combine the URL to a character string
url.call <- paste0(url, "column=", measure.infections, "&row=", time.day, "&row=", region.shp)

# Read the data in to an R object and select the object `COVID-19 cases in the infectious diseases registry`
# Note how this step differs a little from what we did yesterday

data <- fromJSONstat(
  readLines(
    url.call, encoding = "UTF-8"))$`COVID-19 cases in the infectious diseases registry`

# Convert the value column to numeric
data$value <- as.numeric(data$value)
```

#### **2. Data wrangling**

First, we will do some data wrangling to format the data for `ggplot2`. Go through each step carefully to form an understanding of what is happening as you are running through the code. Make sure to copy and run each line!

**2.1** Let's make a copy of the data and format them so that they are suitable for plotting with `ggplot2`.

```{r}
plotdata <- data
```

**2.2** Filter out NA values.

```{r}
plotdata <- plotdata %>% filter(!is.na(value))
```

**2.3** Convert the column containing hospital districts (`hcdmunicipality2020`) into a factor.

```{r}
plotdata$region <- as.factor(plotdata$hcdmunicipality2020)
```

**2.4** Convert the date values from character format to actual dates (this is very useful for plotting).

```{r}
plotdata$week <- as.Date(plotdata$dateweek2020010120201231)
```

**2.5** Check the structure of the data. The output should look like the example below.
```{r}
str(plotdata)
```

```r
'data.frame':	5786 obs. of  6 variables:
 $ dateweek2020010120201231: chr  "2020-01-28" "2020-01-28" "2020-01-28" "2020-01-28" ...
 $ hcdmunicipality2020     : chr  "Åland" "Southwest Finland Hospital District" "Satakunta Hospital District" "Kanta-Häme Hospital District" ...
 $ measure                 : chr  "Number of cases" "Number of cases" "Number of cases" "Number of cases" ...
 $ value                   : num  0 0 0 0 0 0 0 0 0 0 ...
 $ region                  : Factor w/ 22 levels "Åland","All areas",..: 1 21 17 8 16 15 9 18 20 6 ...
 $ week                    : Date, format: "2020-01-28" "2020-01-28" "2020-01-28" "2020-01-28" ...
```

**2.6** Group the data by region to create regional sums (similar to yesterday's excercise).

```r
plotdata_sums <- plotdata %>%
  group_by(region) %>%
  summarise(observations = sum(value, na.rm = TRUE))
```

Ok! Now we're ready to create some plots.

#### **3. Exercises**

**3.1** Plot the regional sums of COVID-19 observations using `plotdata_sums`. Hint: use `geom_col()`.

```r
# Write the answer in your R script

```

**3.2** Now let's create a plot using the original `plotdata` data frame (which has the daily values).
This plot does not look very good...

```r
ggplot(data = plotdata,
       aes(x = week, y = value))+
  geom_line()
```

**3.3** Try improving the previous plot by separating each region. Hint: Add a parameter to `geom_line()`.

```r
# Write the answer in your R script

```

**3.4** Next, let's plot regional cumulative sums of COVID-19 observations. Add a new column that has the daily cumulative sum for each region. Note that we use `cumsum()` to calculate this.

```{r}
plotdata <- plotdata %>%
  group_by(region) %>%
  mutate(cumulative_sum = cumsum(value))
```

**3.5** Plot the cumulative sum. You can try different geoms here.

```{r}
# Write the answer(s) in your R script

```

**3.6** Play around with the plot.

You can, for example, change the y-axis to log scale and select specific regions to plot.

If you are feeling brave, get some more data from the THL website and get creative!

```{r}

# Continue the code below ...

#ggplot(data = ...,
#       aes()) +

```

#### **3. Saving and exporting your script**

If you don't remember how to save and export your data from RStudio Server, follow the instructions set out in the **Starting with Data exercise sheet** (steps 7.3 and 7.4).

That's it in terms of exercises! Depending on the course schedule, there may be a chance to go through the solutions during the beginning of Day 3. However, if you're attending a two-day course feel unsure about how to solve some of the problems above, you can have a look at the completed exercise sheets on eLena.

Thanks for joining the course (so far!) and we hope you've found the contents useful!